---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)

devtools::load_all()

```

# healthdatacsv

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

**healthdatacsv** allows users to query the [healthdata.gov API](https://healthdata.gov/node/25341) catalog and return tidy data frames. This package focuses on the data.json endpoint and will download datasets if available via file download, namely, in csv format.  

## Installation

You can install the development version of **healthdatacsv** from [GitHub](https://github.com/) with:

``` {r, eval = FALSE}
install.packages("remotes")

remotes::install_github("iecastro/healthdatacsv")

```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(healthdatacsv)

# query API based on keywords
fetch_catalog(keyword = "alcohol|drugs")

# query API and dowload data
fetch_catalog("Centers for Disease Control and Prevention",
                               keyword = "built environment") %>% 
  fetch_csv()


```


## Basic workflow

**healthdatacsv** is pipe friendly, so it's easy to integrate into a tidy workflow. For example, say you're interested in searching CDC datasets related to the built environment.  

Let's fetch the catalog of available data products:

```{r}
cdc_built_env <- fetch_catalog("Centers for Disease Control and Prevention",
                               keyword = "built environment")

cdc_built_env

```

In this case, there is only one product available.  To learn more, you can simply `pull` the description:

```{r}
cdc_built_env %>%
  pull(description)

```

This dataset relates to state legislation on nutrition, physical activity, and obesity during 2001-2017.  Data only includes enacted legislation.

To download the data, we pass the catalog df to the `fetch_csv` function.  Since there is only one dataset to download, we can `unnest` in the same pipe.  *If the catalog has more than one product that you'll like to keep, it is recommended to unnest each product separately.* If the catalog consists of several time points of the same dataset, you could unnest in one pipe, given all column names are the same. 


```{r}
data_raw <- cdc_built_env %>%
  fetch_csv() %>% #> 
  tidyr::unnest(data_file)

data_raw %>%
  glimpse()

```


### Some helper functions

`list_agencies()` will query and download names of agencies listed in the catalog. You can enter partial name or initials of agency of interest to check if it has data cataloged in the API. Argument relies on [regular expression](https://stringr.tidyverse.org/articles/regular-expressions.html) to detect string matches.

```{r}

list_agencies("NIH|CDC|FDA")

list_agencies("Institute|Drug")

```

The resulting dataframe will depend on matches to the string supplied. To pull all agencies cataloged, simply use `list_agencies` with the default argument:

```{r}

list_agencies() %>% 
  print(n=Inf)

```

`get_keywords()` will extract all keywords cataloged. The function accepts a full *agency* name (in proper title case) as argument; or it can be run with the default value for a data frame of all keywords and agencies cataloged.

```{r, eval=FALSE}

get_keywords("National Institute on Drug Abuse (NIDA), National Institutes of Health (NIH)")

```



#### Info

Development of this package is partly supported by a research grant from the National Institute on Alcohol Abuse and Alcoholism - NIH Grant #R34AA026745. This product is not endorsed nor certified by either healthdata.gov or NIH/NIAAA.


