---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)

devtools::load_all()

```

# healthdatacsv

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of healthdatacsv is to query the healthdata.gov API and 

## Installation

You can install the development version of **healthdatacsv** from [GitHub](https://github.com/) with:

``` {r, eval = FALSE}
install.packages("remotes")

remotes::install_github("iecastro/healthdatacsv")

```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(healthdatacsv)

# query API based on keywords
fetch_catalog(keyword = "alcohol|drugs")

# query API and dowload data
fetch_catalog("Centers for Disease Control and Prevention",
                               keyword = "built environment") %>% 
  fetch_csv()


```


## Basic workflow

**healthdatacsv** is pipe friendly, so it's easy to integrate into a tidy workflow. For example, say you're interested in searching CDC datasets related to the built environment.  

Let's fetch the catalog of available data products:

```{r}
cdc_built_env <- fetch_catalog("Centers for Disease Control and Prevention",
                               keyword = "built environment")

cdc_built_env

```

In this case, there is only one product available.  To learn more, you can simply `pull` the description:

```{r}
cdc_built_env %>%
  pull(description)

```

This dataset relates to state legislation on nutrition, physical activity, and obesity during 2001-2017.  Data only includes enacted legislation.

To download the data, we pass the catalog df to the `fetch_csv` function.  Since there is only one dataset to download, we can `unnest` in the same pipe.  *If the catalog has more than one product that you'll like to keep, it is recommended to unnest each product separately.* If the catalog consists of several time points of the same dataset, you could unnest in one pipe, given all column names are the same. 


```{r}
data_raw <- cdc_built_env %>%
  fetch_csv() %>% #> 
  tidyr::unnest(data_file)

data_raw %>%
  glimpse()

```

### Info

Development of this package is partly supported by a research grant from the National Institute on Alcohol Abuse and Alcoholism - NIH Grant #R34AA026745. This product is not endorsed nor certified by either healthdata.gov or NIH/NIAAA.


